Do syntactic and semantic similarity lead to interference effects?\\ Evidence from self-paced reading and event-related potentials using German

### 2024-11-08 paper/revision2/ResponseToReviewers.tex

\begin{document}
\hfill \today
%\section{}
%\subsection{}

\noindent \textbf{Prof.\ Jarrold}\\
\noindent \textbf{Associate Editor}\\
\noindent \textbf{Journal of Memory and Language}
\vskip 1em

\noindent \textbf{Response letter for Submission JML-24-17, 2nd round of revisions}
\vskip 2em
\noindent Dear Prof. Jarrold, 
\vskip 1em
Together with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of the article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.
We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.

%add description of revisions here
%To address reviewer \# 2 concern that the detailed description of the SPR results in the critical and surrounding regions might be misleading the reader, we have removed the panels of Figure 3 that highlighted the reading times of these regions. Instead, we now show the long-lasting reading time difference starting at the distractor in each of the self-paced reading sub-experiments.

@@

\subsection*{Comment R1.7}
\begin{quote}
``p. 64: ``Word-by-word presentation is likely to be more demanding on the comprehender's memory, which should make it easier to detect interference effects. So, the differences in methodology do not offer a straightforward
explanation for the lack of syntactic interference in our data compared to the
previous studies."

Actually, I thought here the authors basically provided an explanation of the differences between their findings and the previous findings because of differences in the methodology, even though their last sentence said otherwise. Let me spell out it: since the current methodology (SPR, EEG) is likely more demanding on memory, it is possible that people give up on detailed processing; rather, they do something akin to good-enough processing and do not fully process; consequently, they could still get interference from semantics (since the semantic interference is really simple, it is just lexical semantics and readers don't need to fully parse for that) but they don't correctly parse embedded subjects, hence the syntactic interference is gone. \textbf{The authors in fact come to good-enough processing in Section 7.3, but then it is not linked back to differences between their findings and the other findings (which I think should be, I do think this is a possible explanation).}''
\end{quote}

\subsubsection*{Response to R1.7}
We have added the reviewer's suggested explanation that the employed methods could have especially led to good-enough processing and that this might have contributed to the different findings between studies (see page \pageref{SPR_goodenough}):

\begin{quote}
    \Paste{SPR_goodenough}
\end{quote}

@@

\section*{Reviewer \#2} 
\subsection*{Comment R2.1}
\begin{quote}
``I appreciate the authors' responses to my previous comments. However, despite the revisions, significant weaknesses remain, rendering in my opinion this study primarily a methodological contribution that adds little value to our understanding of interference effects. Additionally, the authors' strong claims often come through as overstatements that could result in serious misconceptions.
My overall assessment is as follows.

1. Interpretability of the SPR results: The results from the self-paced reading (SPR) task remain unclear. The only observable effect—the semantic interference—emerges well before the critical region of interest (specifically, at the distractor region) and persists without change beyond the critical region. As a result, any effect at the critical region is uninterpretable.

The authors acknowledge the interpretability issue with the SPR results in their response to my comment, and they now include the following statement: ``Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over regions, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from further discussing effects occurring in the later regions." \textbf{However, this comes after more than 10 full pages (pp. 25-37) of discussion on the reaction times (RTs) in the critical and pre-critical regions, which would lead any reader to believe these findings are meaningful. If, as the authors now realize, the effect emerges much earlier, this entire discussion is misleading: the SPR results after the distractor region are simply uninterpretable.}''
\end{quote}

\subsubsection*{Response to R2.1}
We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly. We have revised Figure 3 to remove the focus on the critical and surrounding regions (see below). At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects. For comparability with previous findings, we then briefly report the later effects. We have revised the Results section in the following way, staring on page \pageref{revised_SPR}:

\begin{quote}
    \Paste{revised_SPR}
\end{quote}

\setcounter{figure}{2}

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
\end{sidewaysfigure}
\clearpage

We do not share the reviewer's interpretation that the effects in the later regions are meaningless. We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractors. We emphasize the link between the distractor and the pre-critical region by discussing them together in a new section with the title ``Reading times difference prior to retrieval" (starting on page \pageref{prior_to_retrieval}). This new section replaces the former section ``Pre-critical effects".

\begin{quote}
    \Paste{SPR_discussion_revised}
\end{quote}

\textcolor{red}{Figure out how much of the revised text to paste here}

@@

\textcolor{red}{Write something about Himanshu's work in relation to this}

@@

\subsection*{Comment R2.6}
\begin{quote}
``Additional points:

1. Original version: ``By contrast, the syntactic interference effect had only anecdotal evidence under the two narrow priors (Normal-(0, 0.1): BF10 = 2.5, Normal-(0, 0.5): BF10 = 1.2). There was evidence against syntactic interference under wider priors (Normal-(0, 1): BF10 = 0.66, Normal-(0, 5): BF10 = 0.14). What this means is that, in our data, only if we assume a priori that the effect size is relatively small, there is very weak evidence in favor of syntactic interference."

Revised version: ``By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatiotemporal windows (BF10 $<$ 1.2). Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF10 $<$ 1.2)."

\textbf{Did the authors change the analyses? Why what was reported as anecdotal evidence under narrow priors is now reported as no evidence or evidence against?}''
\end{quote}
\subsubsection*{Response to R2.6}
Yes, the priors were changed in the 1st revision (directional priors in the original version and symmetrical priors in the revised version, see also our Response to E.3). This resulted in small numerical changes of the results, e.g., BF10 = 2.5 in the original analysis decreased to BF10 = 1.2 in the revised analysis. Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2. Since the first revision and in line with \cite{jeffreys1998theory} and \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.

### 2024-11-08 paper/revision2/elsarticle-template-harv.tex
Add files via upload

As discussed in detail below, the second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall. The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.

@@

\section{Data availability statement}
{All materials, data and analysis scripts can be accessed via \\ https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}

@@

\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentive reading. The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structures.

{
\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. Panel C shows the pooled reading times of all conditions in the critical and surrounding regions. Panels D and E show the reading times of all conditions in the critical and surrounding regions in the different parts of the self-paced reading experiment.}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom_exp.jpg}
\end{sidewaysfigure}
\clearpage}

It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This pattern persisted in the following regions \textcolor{blue}{-- almost until the end of the sentence.} The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences. To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference. This sustained difference in reading times is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there can not be attributed clearly to the processing of that specific region. However, for comparability with previous work, we briefly report the reading times of the later regions which were identical across conditions, i.e., pre-critical, critical and spill-over region, below.}

Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but \textcolor{blue}{it was only used in the present study.} The pre-pre-critical and pre-critical region showed longer reading times for high vs.\ low semantic interference conditions. The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low. The other two conditions showed intermediate reading times. The spill-over region again primarily showed a difference between high and low semantic interference conditions.

\textcolor{blue}{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b, respectively. Across the sub-experiments, there is some variation in the reading time patterns, but all of them show a pattern consistent with semantic interference in at least some of the regions. Experiment 1a showed a remarkable difference between the experimental sessions, with overall shorter reading times in the second session. This attenuation in reading times was likely due to adaptation to the SPR task.}

\textcolor{blue}{\st{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b. Panel D shows the results from Experiment 1a. The first session of Experiment 1a (items 1-60, N = 204, left sub-figure) showed results similar to the pooled results presented in Panel C. In the second session of Experiment 1a (items 61-120, N = 160, right sub-figure) a semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session. Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times and differences between conditions were likely due to adaptation to the SPR task in the second session. }}

\textcolor{blue}{\st{Panel E shows that Experiment 1b with items 1-60 (N = 350, left sub-figure) showed a pattern consistent with semantic interference in all regions. Experiment 1b with items 61-120 (N = 220, right sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself. At the verb, reading times were longest when both syntactic and semantic interference were high. Reading times were a bit faster when both syntactic and semantic interference were low. The other two conditions (high syntactic/low semantic and low syntactic/high semantic) led to the fastest verb reading times in Experiment 1b with items 61-120. }}

\textcolor{blue}{\st{Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure} \ref{fig:spr_posteriors} \st{shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters. The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions. In the spill-over region, it included only positive values ([1, 15] ms). In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms). The posteriors for the interaction were mostly positive in all regions. The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}

\textcolor{blue}{\st{In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms). The posteriors for the interaction were mostly positive in all regions. The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}

Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions. The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions. The estimates of syntactic interference were less consistent across regions. They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the sustained reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). {There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{ While Bayes factors provided mostly evidence for semantic interference and evidence against syntactic interference and the interaction at the critical and surrounding regions, reading times started to differ earlier in the sentence. Starting at the distractor, high semantic interference conditions (animate distractor) were read slower than low semantic interference conditions (inanimate distractor). Because the distractors differed between conditions, we did not analyze the reading time difference in that region. }

\textcolor{blue}{\st{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region. This means that semantic interference occurred even before the retrieval event could begin. Reading times of the pre-critical word and critical verb showed evidence against syntactic interference. In the spill-over region, the syntactic manipulation slowed down the reading times, but to a lesser extent than the semantic manipulation (CrI of syntactic interference vs. CrI of semantic interference: [1,15] ms vs. [7,21] ms). The Bayes factors also provided less support for the syntactic effect than for the semantic effect (BFs under prior Normal(0, 0.01) in the spill-over region for the syntactic effect vs. the semantic effect: 2.8 vs. 246). The relatively small extent of the syntactic effect which only emerged in one region and which was also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference. However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region. We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion. For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) for a very small effect ($\pm$8 ms). There was evidence against the interaction under wider priors and in the surrounding regions. Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from further discussing effects occurring in the later regions. The effects in the pre-critical region are discussed next.}}

\subsubsection{The effects in the pre-critical region}

At the distractor (which occurred well before the pre-critical region), high semantic interference (animate distractors) led to longer reading times than low semantic interference (inanimate distractors, see Figure \ref{fig:whole_sentence} A and B). One might argue that this pattern could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

\label{plausib_anim_inan}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).

@@

To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility led to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading times slowdown, i.e., semantic interference led to slower reading times independent of plausibility.

\begin{figure}[ht]
    \centering
        \caption{Posteriors of the semantic interference and plausibility effect and their interaction on the reading times in the pre-critical region. The numerical values are the means and 95\,\% credible intervals. The blue vertical lines represent the median and the blue shaded areas are 80\,\% credible intervals.}\label{fig:posteriors_plausibility}
    \includegraphics[width=0.8\linewidth]{posteriors_spr_pooled_774_plausibility.png}
\end{figure}
}

At the pre-critical region, there was evidence for semantic interference. As this cannot be due to retrieval interference, the most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above). The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

@@

\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. They discussed several alternative explanations for their observed effects. {\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.

@@

Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region. However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B). The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase. Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data. Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in later regions as well. The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence. It is not possible to pinpoint whether the effects in later regions were due to a retrieval initiated at the respective region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}. Because of the unclear status of the effects at the critical verb and spill-over region, we refrain from interpreting the semantic interference effects which were observed in these regions.

@@

An obvious difference between the studies is the experimental methodology. The majority of the previous experiments employed eye-tracking while reading, while we used self-paced reading and EEG / ERPs. However, word-by-word self-paced reading (SPR) and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} and RSVP. Word-by-word presentation is likely to be more demanding on the comprehender's memory, which should make it easier to detect interference effects. So, the differences in methodology do not offer a straightforward explanation for the lack of syntactic interference in our data compared to the previous studies.

>>>>

The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall. The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation. For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both). In the Results section, we present the pooled data (N=774) from E1a and E1b.
As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.

@@

\section{Data availability statement}
All materials, data and analysis scripts can be accessed via \\ \href{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}

@@

\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentively performing the task. The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
\end{sidewaysfigure}
\clearpage

\label{revised_SPR}\Copy{revised_SPR}{
It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This pattern \textcolor{blue}{was found in the pooled data (panels A and B) as well as in the data of each experimental session (panels C and D). To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}. Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.

\textcolor{blue}{Crucially, the reading times difference starting at the distractor} persisted in the following regions. The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences. The reading time difference starting at the distractor and persisting almost until the end of the sentence is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there cannot be attributed clearly to the processing of that specific region. However, for comparability with previous work, we briefly report analyses of the reading times of the later regions, i.e., pre-critical, critical and spill-over region, below.}

Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions. The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions. The estimates of syntactic interference were less consistent across regions. They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}
}

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). {There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\subsection{Discussion}
\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{However, reading times started to differ before this critical dependency could be formed. Starting at the distractor which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors). Because the distractor differed between conditions, we did not analyze distractor reading times. Instead, we treat the pre-critical region which was identical across conditions and preceded the critical verb as a proxy to investigate the difference starting at the distractor. Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region. We refrain from interpreting effects which were observed in later regions, i.e., the critical verb and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes. The reading times differences at the distractor and the pre-critical region are discussed below.}}

\subsubsection{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}

\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence} A and B). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

Another possible explanation for the reading time differences starting at the distractor could be \textcolor{blue}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).

@@

To investigate whether this plausibility difference is driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region. This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings. We used the same priors as presented in Table \ref{tab:spr_priors}. The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items. Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes. Although plausibility influenced reading times (higher plausibility led to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms). The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
Normal(0, 0.05): BF$_{10}$ = 2.4,
Normal(0, 0.1): BF$_{10}$ = 1.3). The interaction of semantic interference and plausibility was centered around zero. In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{blue}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions. However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{blue}{high} semantic interference led to slower reading times independent of plausibility.

\begin{figure}[ht]
    \centering
        \caption{Posteriors of the semantic interference and plausibility effect and their interaction on the reading times in the pre-critical region. The numerical values are the means and 95\,\% credible intervals. The blue vertical lines represent the median and the blue shaded areas are 80\,\% credible intervals.}\label{fig:posteriors_plausibility}
    \includegraphics[width=0.8\linewidth]{posteriors_spr_pooled_774_plausibility.png}
\end{figure}

\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference. T}he most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above). The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

@@

\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. {\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.

@@

Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region. However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence}). The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase. Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data. Crucially and as mentioned above, \textcolor{blue}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{blue}{the driver of effects observed in later regions. The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{blue}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.

@@

An obvious difference between the studies is the experimental methodology. The majority of the previous experiments employed eye-tracking while reading, while we used self-paced reading and EEG / ERPs. However, word-by-word self-paced reading (SPR) and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} and RSVP. Word-by-word presentation is likely more demanding on the comprehender's memory, which should make it easier to detect interference effects. So, the differences in methodology do not offer a straightforward
explanation for the lack of syntactic interference in our data compared to the previous studies.  \Copy{SPR_goodenough}{\textcolor{blue}{However, due to the demanding task comprehenders might trade off processing depth and adapt good-enough processing \citep{ferreira2007goodenough}. Consequently, they would not fully parse embedded structures, hence there would be no syntactic interference from embedded subjects. Under this account, semantic interference might still arise because animacy information is readily available from the lexical entries with no need for detailed processing.}}\label{SPR_goodenough}
